### 2014-10-07 paper/include/dtail.tex
more reorganization
stuff
u
wiriten cool stuff on all aproaches, or something.

Zehan Cui et al. \cite{dtail} propses DTail-R in which data for line refresh rate is stored in the DRAM itself instead of storing it in the memory controller which many other solutions do \cite{raidr}\cite{smartrefresh}\cite{refrint}. To counter the added latency of a slower memory they utilizes spatial locality of the refresh data and performs prefetching into a FIFO buffer. The retention time data consists of 4 bits and is interpreted in such way that the refresh period becomes \(64ms \times 2^n, n = [0...7]\). With this structure the amount of memory required is weighing in on a negliable amount of 0.006\% DRAM capacity (i.e. 1.92MB in a 32GB DRAM). One thing that DTail-R misses is refreshrate compensation for temperature variations. As the refresh rate is set per line DTail-R gets a higher resolution than techniques that set refreshrate on a higher level. Therefore DTail-R manages to decrease the amount of refreshes with as much as 87.9\% and thereby only have 16.7\% energy overhead compared to a refresh less 32GB DRAM. Zehan Cui et al. also proposes DTail-V \cite{dtail} which can be combined with DTail-R and is further discussed in \ref{sec:val}.

>>>>

Z. Cui et al. \cite{dtail} proposes DTail-R in which data for line refresh rate is stored in the DRAM itself instead of storing it in the memory controller which many other solutions do \cite{raidr}\cite{smartrefresh}\cite{refrint}. To counter the added latency of a slower memory they utilizes spatial locality of the refresh data and performs prefetching into a FIFO buffer. The retention time data consists of 4 bits and is interpreted in such way that the refresh period becomes \(64ms \times 2^n, n = [0...7]\). With this structure the amount of memory required is weighing in on a negligible amount of 0.006\% DRAM capacity (i.e. 1.92MB in a 32GB DRAM). One thing that DTail-R misses is refresh rate compensation for temperature variations. As the refresh rate is set per line DTail-R gets a higher resolution than techniques that set refresh rate on a higher level. Therefore DTail-R manages to decrease the amount of refreshes with as much as 87.9\% and thereby only have 16.7\% energy overhead compared to a refresh less 32GB DRAM.

Another technique proposed by Z. Cui et al. is DTail-V \cite{dtail} which can be combined with DTail-R. DTail-V uses \textbf{V} data to mitigate unnecessary refreshes.

### 2014-10-07 paper/include/reducing_refresh_power.tex
wiriten cool stuff on all aproaches, or something.

\subsubsection*{\textbf{Retention Time}}
.

\todo[inline]{Replace the copy-paste texts}
Retention time refers to the period during which cells hold valid values as charge gradually leaks. Process variation [10][16] causes the retention time of DRAM cells to vary across the chip. Note that approaches that ex-ploit R information are insensitive to system workloads and global memory usage, which makes them attractive omponents for a refresh-optimized memory subsystem.

\subsubsection*{\textbf{Data Tolerance}}
.

\todo[inline]{Replace the copy-paste texts}
Applications like games, media processing, machine learning, and unstructured information analysis tolerate errors in portions of their data and still produce acceptably accurate results. Approximate computation [31, 8] exploits this approximate data to realize tradeoffs among performance, energy, and accuracy. Cells containing such error-tolerant T data need not be refreshed as often as those containing critical data. How many cells fall into this category depends on application characteristics.

\subsubsection*{\textbf{Validity}}
.

\todo[inline]{Replace the copy-paste texts}
If the OS has not allocated the page frame, its data are meaningless, and refreshes to it are wasteful. Severa
l software approaches thus attempt to trigger refreshes only for rows with valid data. The effectiveness of schemes using such V refresh information is sensitive to the total memory usage of the system.

>>>>

\subsubsection*{\textbf{Retention Time}}
The need for refresh is as mentioned caused by leakage in the memory cells. Retention time is a measurement of the time it takes before a cell loose its data if not refreshed (recharged). Variations in the manufacturing process causes fluctuation in the retention time of the DRAM cells on the same chip. To acomodate for the cells with the lowest retention time manufaturers produce chips with average retention times much higher than possible minimum. This gives room for optimications using variable refresh rate for different cells. The retention time is also highly sensitive to temperature variations and becomes drastically lower at higher temperatures. For techniques using this aproach the biggest advatage is that the impovments are independent of workload, except for the potential rise in temperature.

\subsubsection*{\textbf{Data Tolerance}}
In the most aproaches all data loss is seen as intolerable. In this aproach data is categorised depending on how critical it is and e.g. aproximate computations, games, or media processing can tollerate some bit errors while still producing a acceptable result. This yields that cells containing data from such applications may be refreshed more seldom. Using this aproach the energy and performance savings achieved is highly depending on what kind of applications is most commonly run on the system.

% Applications like games, media processing, machine learning, and unstructured information analysis tolerate errors in portions of their data and still produce acceptably accurate results. Approximate computation [31, 8] exploits this approximate data to realize tradeoffs among performance, energy, and accuracy. Cells containing such error-tolerant T data need not be refreshed as often as those containing critical data. How many cells fall into this category depends on application characteristics.

\subsubsection*{\textbf{Validity}}
In a system that does not uitilze all the availible memory space it is unnesessary to refresh parts not used. Therefore, cells containing data that are not valid does not need to be refreshed. This aproach is highly sensitive to memory uitilization of the system and its efects on energy consumption and performance improvments will be diminishing on systems with high workload.

%If the OS has not allocated the page frame, its data are meaningless, and refreshes to it are wasteful. Several software approaches thus attempt to trigger refreshes only for rows with valid data. The effectiveness of schemes using such V refresh information is sensitive to the total memory usage of the system.
